---
title: "Description of supported items"
author: "Jianfeng Li"
date: "`r Sys.Date()`"
output:
  prettydoc::html_pretty:
    toc: true
    theme: cayman
    highlight: github
  pdf_document:
    toc: true
vignette: >
  %\VignetteIndexEntry{Description of supported items}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
  %\VignetteEncoding{UTF-8}
---

```{r, echo = FALSE}
knitr::opts_chunk$set(comment = "#>", collapse = TRUE, screenshot.force = FALSE)
get_dt <- function(item) {
  Name <- names(item)
  Description <- unname(unlist(lapply(item, function(x){
    if (is.na(x["description"])) return("")
    return(x["description"])}))
  )
  Publication <- unname(unlist(lapply(item, function(x){
    if (is.na(x["publication"])) return("")
    return(x["publication"])}))
  )
  data.frame(Name, Description)
}
```

## Description of supported items

Partial meta information of tools/scripts and databases are saved in the followed files. 
You can get the file list by command `unname(unlist(get.meta()[1:4]))`.

We do not show the meta information of [annovarR](https://github.com/JhuangLab/BioInstaller/blob/master/inst/extdata/config/db/db_annovar.toml) supported items in this vignette.

```{r}
library(BioInstaller)
meta_info <- get.meta(read.config.params = list(rcmd.parse = TRUE))
unname(unlist(meta_info[1:4]))
```

The followed tables show partial items description that have been collected in BioInstaller package including `Github items`, `Non-github items` and `Database items`.

### Github items

```{r echo = FALSE}
DT::datatable(get_dt(meta_info$github$item), escape = FALSE)
```

### Non-github items

```{r echo = FALSE}
DT::datatable(get_dt(meta_info$nongithub$item), escape = FALSE)
```

### Database items

```{r echo = FALSE}
DT::datatable(get_dt(meta_info$db$item), escape = FALSE)
```

## A curated list of resources for learning bioinformatics

BioInstaller also provides a simplified and curated list of resources for learning bioinformatics, such as sequencing method, web service, command-line tools and database resources. You can get it from [here](https://github.com/JhuangLab/Bioinformatics-Resources).

