---
title: "Examples of Templet Configuration File"
author: "Jianfeng Li"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Examples of Templet Configuration File}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r, echo = FALSE}
knitr::opts_chunk$set(comment = "#>", collapse = TRUE)
```

Templet configuration files in BioInstaller are important. All of softwares and databases source URL, install script and other information were stored in these configuration files. 

Most of configuration files be parsed by [configr](https://github.com/Miachol/configr). Compared with original configr package `#R# R CMD #R#` is a different point which can be used to mark those R format command. 

## github.toml and nongithub.toml

Built-in configuration files: `github.toml` and `nongithub.toml` can be used to download and install several software and database. `install.bioinfo(show.all.names = TRUE)` can be used to get all of avaliable softwares and databases existed in github.toml and nongithub.toml.

### Github Softwares

Some of item in github configuration file can be used to control the BioInstaller behavior:

- `use_git2r` be setted to `false`, BioInstaller will use the [git](https://en.wikipedia.org/wiki/Git) of your system.
- `use_git2r` be setted to `false` and `recursive_clone` be setted to `true`, BioInstaller will do that `git clone --recursive https://path/repo`
- `before_install` can be used to finish some pre-process step
- `install` is the mainly item to control the install step. Moreover, you can use your install script saved in a file and run it by `#R# system('/path/yourscript')#R#`
- `make_dir` is the compile directory of softwares which depend on your install script, workdir of R default be setted to `download.dir` and will be changed to `make_dir` when running install step.
- `bin_dir` will be saved when finish install step

```toml
[bwa]
github_url = "https://github.com/lh3/bwa"
after_failure = "echo 'fail!'"
after_success = "echo 'successful!'"
make_dir = ["./"]
bin_dir = ["./"]

[bwa.before_install]
linux = ""
mac = ""

[bwa.install]
linux = "make"
mac = "make"
```

Github softwares version control can be done by `git2r` package and github tag [API](https://developer.github.com/v3). Source url of softwares or files deposited in github can be found by `github_url` in `github.toml`.

### Non-Github Softwares or Databases

Configuration file of non-github softwares and databases are similar to githubs:
- `github_url` be replaced by `source_url` 
- `url_all_download` be setted to `true` if you want to download mulitple files. 
- [rvest](https://cran.r-project.org/package=rvest) and [RCurl](https://cran.r-project.org/package=RCurl) packages can be used to parse the version infomation of non-github softwares or databases.
- `version_order_fixed` can be setted to `true` if you don't want to use the built-in version reorder function. If you set `url_all_download` to `false`, which can let us using multiple mirror to avoid one of invalid URL.

```toml
[gmap]
# {{version}} will be parsed to your install.bioinfo `version` parameter
# or the newest version parsed from fetched data.
source_url = "http://research-pub.gene.com/gmap/src/{{version}}.tar.gz"
after_failure = "echo 'fail!'"
after_success = "echo 'successful!'"
make_dir = ["./"]
bin_dir = ["./"]

[gmap.before_install]
linux = ""
mac = ""

[gmap.install]
linux = "./configure --prefix=`pwd` && make && make install"
mac = ["sed -i s/\"## CFLAGS='-O3 -m64' .*\"/\"CFLAGS='-O3 -m64'\"/ config.site",
"./configure --prefix=`pwd` && make && make install"]
```

Version control of non-github softwares and databases need a function parsing URL and use `{{version}}` to replace in the `source_url`.

## nongithub_databases_blast.toml

This configuration file can be used to download NCBI blast database: `install.bioinfo(nongithub.cfg = system.file('extdata', 'databases/blast.toml', package = 'BioInstaller'), show.all.names = TRUE)`.

BioInstaller use [configr](https://github.com/Miachol/configr) `glue` to reduce the length of files name. That using less word to storage more files name. More useful databases FTP url can be accessed in the future. I hope you can set your own configuration file not only use the BioInstaller built-in configuration files.

```{r}
library(configr)
library(BioInstaller)
blast.databases <- system.file('extdata', 
  'databases/blast.toml', package = 'BioInstaller')

read.config(blast.databases)$blast_db_nr$source_url
read.config(blast.databases, glue.parse = TRUE)$blast_db_nr$source_url
mask.github <- tempfile()
file.create(mask.github)
install.bioinfo(nongithub.cfg = blast.databases, github.cfg = mask.github,
  show.all.names = TRUE)
```

## Automatic parse from BIO_SOFTWARES_DB_ACTIVE database
To resolve the dependence and its path, BioInstall can automatic recogniztion the `{{key:value}}` format expression and get its real value from BioInstall `BIO_SOFWARES_DB_ACTIVE` database which were be setted by environment variable BIO_SOFTWARES_DB_ACTIVE and parameter `db`.

For example, Pindel need the `htslib` to finish its install, and we use `./INSTALL {{htslib:source.dir}}` as the install step of Pindel that the value be setted in `system.file("extdata", "github.toml", package = "BioInstaller")` file Pindel section. In R, the `{{htslib:source.dir}}` will be replaced by the value that were stored in BIO_SOFTWARES_DB_ACTIVE or `db`, a parameter of `install.bioinfo` function, which indicating the BioInstall Database PATH. More example can be founded in [configr](https://github.com/Miachol/configr) `parse.extra` 'other.config' parameter.

## Automatic parse from install.bioinfo parameter extra.list
To improve the flexibility of configuration templet, BioInstall can automatic recogniztion the `{{parameters}}` format expression and get its value if its existed in `install.bioinfo` extra.list parameter and `name`, `version`, `os.version`, `destdir` were default be setted.

For example, GMAP source_url need to point the `version` value, and we use `source_url = "http://research-pub.gene.com/gmap/src/{{version}}.tar.gz"` as the download URL and can be used to install different version GMAP that the value be setted in `system.file("extdata", "nongithub.toml", package = "BioInstaller")` file gmap section. In R, the `{{version}}` will be replaced by the `version` parameter value of `install.bioinfo` (if version were `NULL`, it will be setted to be the newst version). More example can be founded in [configr](https://github.com/Miachol/configr) `parse.extra` 'extra.list' parameter.

## Automatic parse from R CMD
For example, `@>@ str_replace('{{version}}', '-linux64', '') @<@`. If this expression `{{version}}` be parsed to `1.2.0-linux64` in configuration file, the full expression will be parsed to `1.2.0`. So, you can write yourself install scripts or config file conveniently. More example can be founded in [configr](https://github.com/Miachol/configr) 'rcmd.parse' parameter.
